SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-07 results.tex
done

This section discusses the results obtained from running PPF in terms of
prefetch cache coverage and speedup. For the SPEC CPU 2017 benchmarks, first
we present the results for single-threaded workloads then for multi-core
workloads.

@@

% djimenez: why is this figure described this way? it's not OK to just
% highlight noticeable improvement. the benchmarks must be selected by some
% unbiased criteria, e.g. some minimum MPKI under LRU or some minimum speedup
% under a baseline prefetcher that isn't PPF.

Figure~\ref{Fig:SPEC2017_1core} shows the IPC speedup obtained by PPF,
compared to BOP, DA-AMPM and SPP; for the individual SPEC CPU 2017 
applications, followed by memory intensive subset and finally the full
suite. All the results have been normalized to the baseline of no prefetching.

@@

PPF yields a geometric mean speedup of \textbf{26.95\%} over the baseline. 
This is equivalent to \textbf{4.63\%} over DA-AMPM, \textbf{4.61\%} over BOP 
and \textbf{3.78\%} over SPP. Out of the 95 simpoints developed for SPEC 
CPU 2017, PPF nearly matches or outperforms most of the prefetchers on 91 
simpoints. 
An interesting case to see here is {\tt 623.xalancbmk}. Despite SPP 
underperforming on that application, PPF manages to considerably outperform 
the next best prefetcher -- BOP. Since this benchmark has varying prefetch 
deltas, SPP's conservative throttling mechanism catches that and quickly halts 
prefetching at an average depth of 2.1. On the other hand, PPF's more efficient 
accuracy check enables it to prefetch upto a lookahead depth of 3.3. Doing this, PPF 
suggests 1.61 times more total prefetches and 2.53 times more 
useful prefetches than SPP.

The only benchmark where PPF fails to match the improvement offered by other prefetchers 
is {\tt 607.cactuBSSN\_s}. \newline Based on our observation of prefetching behavior, we gather that
BOP's aggressive and localised nature can tackle this workload really well; as opposed
to SPPs lookahead nature. As a result, SPP, and hence PPF, underperform on this benchmark.

At its peak, PPF manages a speedup of a factor of over \textbf{3.39x} on one of the
{\tt 602.gcc\_s} traces. This also corresponds to speedup gain of
\textbf{60\%} over the next best prefetcher for that trace -- BOP. In general, benchmarks
{\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt {623.xalancbmk}\_s}~~and~~{\tt
649.fotonik3d\_s}  benefit the most from PPF, with the speedup over SPP ranging
from \textbf{10\% to 25\%}.

% djimenez: is this geometric mean? if so, say so.
% [EB] Resolved
On the full SPEC CPU 2017 suite, PPF improves the geometric mean IPC of 
the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better 
than the next best prefetcher -- SPP.\\ \\
%
\begin{figure}[ht]
\includegraphics[width=\columnwidth]{Coverage}
\caption{Fraction of Cache Misses Covered}
\label{Fig:Coverage}
\end{figure}
%
% djimenez: this is a weird way to make a header for a section. use \paragraph or \subsection
\noindent \textbf{Coverage:}
Prefetcher coverage is defined as the ratio of the number of misses avoided
through prefetching to the number of misses with no prefetching.

Figure~\ref{Fig:Coverage} shows the fraction of misses in the L2 and LLC
avoided by the various prefetchers. PPF has the highest coverage of all the
prefetchers simulated. On the SPEC CPU 2017 benchmarks, PPF reduces misses by
\textbf{75.5\%} and \textbf{86.9\%} in the L2 and LLC respectively. For the
same benchmark, the next best prefetcher, DA-AMPM, covers \textbf{54.3\%} and
\textbf{78.5\%} of the misses respectively.

@@

\subsection{Multi-core Results}
\label{Results-Multi}
In this section, we demonstrate the improvement achieved by PPF for a mix of
multi-programmed workloads.\\ \\
%
\noindent \textbf{4-Core Environment:}
Figure~\ref{Fig:4Core_SPEC2017} shows a comparison of speedups obtained on
4-core mixes of a memory intensive subset SPEC CPU 2017. We plot all 4
prefetchers, normalized to the baseline. The workloads have been sorted in
increasing order of the speedup. PPF offers a speedup of \textbf{51.2\%} on
these traces, an improvement of \textbf{11.4\%} over the underlying SPP,
\textbf{9.7\%} over the next DA-AMPM, and \textbf{16.9\%} over BOP.
On a different set of fully random SPEC CPU 2017 4-core mixes (not illustrated
for space reasons), PPF provides an IPC speedup of \textbf{26.07}\% over the
baseline, which is an improvement of \textbf{5.6}\% over SPP.\\ \\
%
\begin{figure}[ht]
\includegraphics[width=2.75in]{8Core_SPEC2017}
\caption{Speedup for 8-core SPEC CPU 2017}
\label{Fig:8Core_SPEC2017}
\end{figure}
%
\noindent \textbf{8-Core Environment:}
The sorted comparison of speedups on the memory intensive 8-core mixes is
shown in Figure~\ref{Fig:8Core_SPEC2017}. PPF improves baseline performance
by \textbf{37.6\%}, an improvement of \textbf{9.7\%} over SPP. For a random
set of SPEC CPU 2017 mixes (not illustrated for space reasons), PPF improves
performance by \textbf{23.4\%} over the baseline, corresponding to
\textbf{4.6\%} over SPP. This increased improvement achieved by PPF over the
base engine SPP in a multi-core environment is expected as PPF has a very
accurate filter, it eliminates useless prefetches before they can cause
pollution in the shared LLC.
BOP offers a better improvement than SPP for the memory intensive mixes. This
superiority can be attributed to BOP's inherent aggressive nature. DA-AMPM is
also ahead of SPP in both the mixes. Interestingly, in all these cases, PPF
consistently outperforms the best performing prefetcher.

@@

\subsection{Additional Memory Constraints}
\label{Results-AdditionalMem}

%Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
%followed by the overall performance on the complete benchmark.  
%For the sake of brevity, we have only shown selected traces

We also model PPF with reduced LLC and with low bandwidth constraints,
respectively (not illustrated for space reasons). Benchmark {\tt 605.mcf\_s}
in low bandwidth conditions is prefetch averse. In general, any prefetcher
yields a negative speedup on that trace. On {\tt 654.roms\_s} and {\tt
607.cactuBSSN\_s}, PPF is unable to match the performance achieved by the best
prefetcher. On the other hand, PPF outperforms all the other prefetchers on
{\tt 623.xalancbmk\_s} and {\tt 638.imagick\_s} benchmarks. Overall, PPF
gives a better improvement under small LLC condition and matches the best
prefetcher, BOP, under low DRAM bandwidth conditions.

>>>>

This section discusses the results obtained from running PPF in terms
of speedup and prefetch cache, for the SPEC CPU 2017 benchmarks.
First, we present the results for single-threaded workloads then for
multi-core workloads.

@@

% djimenez: why is this figure described this way? it's not OK to just
% highlight noticeable improvement. the benchmarks must be selected by
% some unbiased criteria, e.g. some minimum MPKI under LRU or some
% minimum speedup under a baseline prefetcher that isn't PPF.

Figure~\ref{Fig:SPEC2017_1core} shows the single core speedup obtained
by BOP, DA-AMPM, SPP and PPF for each of the individual SPEC CPU 2017
applications, followed by the geomean of the memory intensive subset
and finally the geomean across the full suite.  All the results have
been normalized to the baseline of no prefetching.

@@

PPF yields a geometric mean speedup of \textbf{26.95\%} over the
baseline.  This is equivalent to \textbf{4.63\%} over DA-AMPM,
\textbf{4.61\%} over BOP and \textbf{3.78\%} over SPP.  Out of the 95
simpoints developed for SPEC CPU 2017, PPF nearly matches or
outperforms most of the prefetchers on 91 simpoints.  One interesting
case here is {\tt 623.xalancbmk}.  Despite SPP under performing on that
application, PPF manages to considerably outperform all prefetchers.
Since this benchmark has varying prefetch deltas, SPP's conservative
throttling mechanism catches that and quickly halts prefetching at an
average depth of 2.1.  On the other hand, PPF's more efficient
accuracy check enables it to prefetch upto a lookahead depth of
3.3. Doing this, PPF suggests 1.61 times more total prefetches and
2.53 times more useful prefetches than SPP.

The only benchmark where PPF fails to match the improvement offered by
other prefetchers is {\tt 607.cactuBSSN\_s}.  Based on our observation
of prefetching behavior, we gather that BOP's aggressive and localized
nature fits this workload very well; as opposed to SPPs lookahead
nature.  As a result, SPP, and hence PPF, underperform on this
benchmark.

\pg{is the following text old? doesn't look right:} At its peak, PPF
manages a speedup of a factor of over \textbf{3.39x} on one of the
{\tt 602.gcc\_s} traces. This also corresponds to speedup gain of
\textbf{60\%} over the next best prefetcher for that trace -- BOP. In
general, benchmarks {\tt 603.bwaves\_s}, {\tt 605.mcf\_s}, {\tt
  {623.xalancbmk}\_s}~~and~~{\tt 649.fotonik3d\_s} benefit the most
from PPF, with the speedup over SPP ranging from \textbf{10\% to
  25\%}.

% djimenez: is this geometric mean? if so, say so.
% [EB] Resolved
On the full SPEC CPU 2017 suite, PPF improves the geometric mean IPC
of the baseline by \textbf{15.24\%}, which is \textbf{2.27\%} better
than the next best prefetcher -- SPP.

\begin{figure}[ht]
\includegraphics[width=\columnwidth]{Coverage}
\caption{Fraction of Cache Misses Covered}
\label{Fig:Coverage}
\end{figure}
%
% djimenez: this is a weird way to make a header for a section. use \paragraph or \subsection

\noindent \textbf{Coverage:} Prefetcher coverage is defined as the
ratio of the number of misses avoided through prefetching over the
number of misses with no prefetching.  Figure~\ref{Fig:Coverage} shows
the fraction of misses in the L2 and LLC avoided by the various
prefetchers.  PPF has the highest coverage of all the prefetchers
simulated.  On the SPEC CPU 2017 benchmarks, PPF reduces misses by
\textbf{75.5\%} and \textbf{86.9\%} in the L2 and LLC
respectively. For the same benchmarks, the next best prefetcher,
DA-AMPM, covers \textbf{54.3\%} and \textbf{78.5\%} of the misses
respectively.

@@

\subsection{Multi-core Results}
\label{Results-Multi}
In this section, we demonstrate the improvement achieved by PPF for a
mix of multi-programmed workloads.

\noindent \textbf{4-Core:} Figure~\ref{Fig:4Core_SPEC2017}
shows a comparison of speedups obtained on 4-core mixes of a memory
intensive subset SPEC CPU 2017. We plot all 4 prefetchers, normalized
to the baseline. The workloads have been sorted in increasing order of
the speedup. PPF offers a speedup of \textbf{51.2\%} on these traces,
an improvement of \textbf{11.4\%} over the underlying SPP,
\textbf{9.7\%} over the next DA-AMPM, and \textbf{16.9\%} over BOP.
On a different set of fully random SPEC CPU 2017 4-core mixes (not
illustrated for space reasons), PPF provides an IPC speedup of
\textbf{26.07}\% over the
baseline, which is an improvement of \textbf{5.6}\% over SPP.

\begin{figure}[ht]
\includegraphics[width=2.75in]{8Core_SPEC2017}
\caption{Speedup for 8-core SPEC CPU 2017}
\label{Fig:8Core_SPEC2017}
\end{figure}

\noindent \textbf{8-Core:} The sorted comparison of speedups on the
memory intensive 8-core mixes is shown in
Figure~\ref{Fig:8Core_SPEC2017}.  PPF improves baseline performance by
\textbf{37.6\%}, an improvement of \textbf{9.7\%} over SPP. For a
random set of SPEC CPU 2017 mixes (not illustrated for space reasons),
PPF improves performance by \textbf{23.4\%} over the baseline,
corresponding to \textbf{4.6\%} over SPP. This increased improvement
achieved by PPF over the base engine SPP in a multi-core environment
is expected as PPF is a very accurate filter. Thus, it eliminates
useless prefetches before they can cause pollution in the shared LLC.
BOP offers a better improvement than SPP for the memory intensive
mixes. This superiority can be attributed to BOP's inherent aggressive
nature. DA-AMPM is also ahead of SPP in both the mixes. Interestingly,
in all these cases, PPF consistently outperforms the best performing
prefetcher.

@@

% \subsection{Additional Memory Constraints}
% \label{Results-AdditionalMem}

% %Figure~\ref{Fig:AddnConstr}(a) and (b) show the performance of PPF in a
% %followed by the overall performance on the complete benchmark.  
% %For the sake of brevity, we have only shown selected traces

% We also model PPF with reduced LLC and with low bandwidth constraints,
% respectively (not illustrated for space reasons).  % Benchmark {\tt
% % 605.mcf\_s} in low bandwidth conditions is prefetch averse. In
% % general, any prefetcher yields a negative speedup on that trace. On
% % {\tt 654.roms\_s} and {\tt 607.cactuBSSN\_s}, PPF is unable to match
% % the performance achieved by the best prefetcher. On the other hand,
% % PPF outperforms all the other prefetchers on {\tt 623.xalancbmk\_s}
% % and {\tt 638.imagick\_s} benchmarks.
% Overall, PPF provides a greater improvement under small LLC condition
% and matches the best prefetcher, BOP, under low DRAM bandwidth
% conditions.
